---
title: "R Notebook"
output: html_notebook
---

```{r}
setwd('C:/Users/Samay Panwar/Documents/GitHub/MH3511-DataAnalysis')
#install.packages("wooldridge")
#install.packages("lmtest")
#install.packages('regclass')
#install.packages('gclus')

library(wooldridge)
library(dplyr)
library(lmtest)
library(regclass)
library(tidyverse)
library(gclus)
```
```{r}
data('bwght')

write.csv(bwght, "C:/Users/Samay Panwar/Documents/GitHub/MH3511-DataAnalysis/bwght.csv")
bw <- na.omit(bwght)
```


```{r}
bw
```

```{r}
cleanedData <- select(bw, -c("packs", "bwghtlbs", "lbwght", "lfaminc", "cigtax", "cigprice"))
cleanedData$male = as.factor(cleanedData$male)
cleanedData$white = as.factor(cleanedData$white)
cleanedData$bwght = as.numeric(cleanedData$bwght)
cleanedData$faminc = as.numeric(cleanedData$faminc)
cleanedData = na.omit(cleanedData)

str(cleanedData)
```
```{r}
# Make a correlation scatterplot with lowess lines
categoricalVariables = c('male', 'white')
correlationData = select(cleanedData, -all_of(categoricalVariables))
pairs(~bwght + faminc + fatheduc + motheduc + parity ,data=correlationData,
   main="Scatterplot Matrix")
```

# Testing for Correlations with the Numerical Variables

```{r, fig.width=8, fig.height=8}

correlationData.r <- abs(cor(correlationData)) # get correlations
correlationData.colors <- dmat.color(correlationData.r) # get colors
# reorder variables so those with highest correlation
# are closest to the diagonal
correlationData.ordered <- order.single(correlationData.r)
cpairs(correlationData, correlationData.ordered, panel.colors=correlationData.colors, gap=.1, 
main="Scatterplot Matrix" )
```
# ANOVA for all Categorical Variables

```{r}
# Conduct ANOVA for all categorical variables
par(mfrow=c(1,2))
boxplot(bwght~male, data = cleanedData, main='Comparison of Baby weight means by "Male"', cex.main=0.7)
boxplot(bwght~white, data = cleanedData, main='Comparison of Baby weight means by "White"', cex.main=0.7)
```


```{r}
male.aov = aov(bwght ~ male, data=cleanedData)
summary(male.aov)
```
```{r}
white.aov = aov(bwght ~ white, data=cleanedData)
summary(white.aov)
```
```{r}
boxplot(bwght~parity, data = cleanedData, main='Comparison of Baby weight means by "Parity"')
```


```{r}
parity.aov = aov(bwght ~ parity, data=cleanedData)
summary(parity.aov)
```
# Conducting Simple Linear Regression for all the Independent Variables

```{r}
# Conduct simple regression models for all remaining variables

SLRObjects = lapply( select(cleanedData, -c('bwght')), function(x) lm(cleanedData$bwght ~ x))

```

# Conducting Multiple Linear Regression

```{r}
# Conduct Multiple Linear Regression with all variables

linearRegression <- lm(bwght~., data=cleanedData)
summary(linearRegression)

```
# Checking for Variance Inflation Factor and Heteroskedasticity
```{r}
# Check for VIF and Heteroskedasticity

bptest(linearRegression)
VIF(linearRegression)

```
# QQ-plots of all the Regressions run

```{r, fig.width=5, fig.height=25}
# Make qq plots of all residuals for all the regressions conducted
par(mfrow=c(7,1))

qqnorm(resid(SLRObjects$faminc), main = 'Residual QQ-plot for Family Income')
qqline(resid(SLRObjects$faminc), col='red')

qqnorm(resid(SLRObjects$fatheduc), main = 'Residual QQ-plot for Father Education')
qqline(resid(SLRObjects$fatheduc), col='red')

qqnorm(resid(SLRObjects$motheduc), main = 'Residual QQ-plot for Mother Education')
qqline(resid(SLRObjects$motheduc), col='red')

qqnorm(resid(SLRObjects$parity), main = 'Residual QQ-plot for Parity')
qqline(resid(SLRObjects$parity), col='red')

qqnorm(resid(SLRObjects$male), main = 'Residual QQ-plot for Male')
qqline(resid(SLRObjects$male), col='red')

qqnorm(resid(SLRObjects$white), main = 'Residual QQ-plot for White')
qqline(resid(SLRObjects$white), col='red')

qqnorm(resid(SLRObjects$cigs), main = 'Residual QQ-plot for Cigs')
qqline(resid(SLRObjects$cigs), col='red')

```
```{r}
qqnorm(resid(linearRegression), main = 'Residual QQ-plot for MLR')
qqline(resid(linearRegression), col='red')
```

# Explain Beta Coefficients for all the regressions

when the family income increases by 1 unit then, on average, the birth weight of the baby increases by 0.08965 grams, ceteris paribus

when the father's education increases by 1 unit then, on average, the birth weight of the baby increases by 0.6102 grams, ceteris paribus

when the mother's education increases by 1 unit then, on average, the birth weight of the baby increases by 0.3755 grams, ceteris paribus

when the number of children already born increases by 1 unit then, on average, the birth weight of the baby increases by 1.601 grams, ceteris paribus

when the child's sex is male, on average, the birth weight of the baby increases by 3.759 grams, ceteris paribus

when the child's race is white then, on average, the birth weight of the baby increases by 4.865 grams, ceteris paribus

when the number of cigs smoked by the mother while pregnant increases by 1 unit then, on average, the birth weight of the baby increases by -0.6203 grams, ceteris paribus

```

